Modelling last glacial cycle ice dynamics in the Alps

### 2017-07-07 articles/cycle/alps-cycle.tex
Rework ice rheology subsection.

@@

      Ice deformation is governed by the constitutive law for ice
      \citep{Glen.1952, Nye.1953},
%
\begin{align}
&\label{eqn:glenslaw}
&\vec{\dot{\epsilon}} = A\,\tau_{\mathrm{e}}^{n-1}\,\vec{\tau} \,.
\end{align}
%
      where $\vec{\dot{\epsilon}}$ is the strain-rate tensor,
      $\vec{\tau}$ the deviatoric stress tensor, and $\tau_{\mathrm{e}}$ the
      effective stress defined in our case by ${\tau_{\mathrm{e}}}^2 = \frac{1}{2}
      \mathrm{tr}(\vec{\tau}^2)$. The ice softness coefficient, $A$, depends
      on ice temperature, $T$, pressure, $p$, and water content, $\omega$,
      through a~piece-wise Arrhenius-type law,
%
\begin{align}
&\label{eqn:softness}
&A = E\cdot
\begin{cases}
A_{\mathrm{c}} \,e^\frac{-Q_{\mathrm{c}}}{RT_{\text{pa}}}            & \text{if}\ T_{\text{pa}}  <  T_{\mathrm{c}} \,, \\
A_{\mathrm{w}} (1+f\omega)\,e^\frac{-Q_{\mathrm{w}}}{RT_{\text{pa}}} & \text{if}\ T_{\text{pa}} \ge T_{\mathrm{c}} \,,
\end{cases}
\end{align}
%
      where $T_{\text{pa}}$ is the pressure-adjusted ice temperature
      calculated using the Clapeyron relation, ${T_{\text{pa}} = T - \beta
      p}$. $R=8.31441$\,\unit{J\,mol^{-1}\,K^{-1}} is the ideal gas
      constant, and $A_{\mathrm{c}}$, $A_{\mathrm{w}}$, $Q_{\mathrm{c}}$ and
      $Q_{\mathrm{w}}$, are constant parameters corresponding to values
      measured below and above a~critical temperature threshold
      $T_{\mathrm{c}}=-10$\,\unit{{\degree}C}
      \citep[p.~72]{Paterson.Budd.1982,Cuffey.Paterson.2010}. The water
      fraction, $\omega$, is capped at a~maximum value of 0.01, above which
      no measurements are available \citep[Eq.~5.7]{Lliboutry.Duval.1985,
      Greve.1997}. Finally, $E$ is a~non-dimensional enhancement factor
      which can take different values, $E_{\text{SIA}}$, in the SIA and
      $E_{\text{SSA}}$, in the SSA.

      In all our simulations, we set constant the power-law exponent, $n=3$,
      according to \citet[p.~55--57]{Cuffey.Paterson.2010}, the Clapeyron
      constant, $\beta=7.9\times 10^{-8}$\,\unit{K\,Pa^{-1}}, according to
      \citet{Luthi.etal.2002}, the water fraction coefficient, $f=181.25$,
      according to \citet{Lliboutry.Duval.1985}, and the SSA enhancement
      factor, $E_{\text{SSA}}=1$, according to
      \citet[p.~77]{Cuffey.Paterson.2010}. These fixed parameter values are
      summarized in Table~\ref{tab:params}.

      On the other hand, we test the model sensitivity
      (Sect.~\ref{sec:sens}) to different values for the two creep
      parameters, $A_{\mathrm{c}}$ and $A_{\mathrm{w}}$, the two activation
      energies, $Q_{\mathrm{c}}$ and $Q_{\mathrm{w}}$, and the SIA
      enhancement factor, $E_{\text{SIA}}$, as follows.
%
\begin{itemize}
  \item Our \emph{default} configuration used in the control run of the
    sensitivity study and all other simulations in the paper includes
    rheological parameters,
    $A_{\mathrm{c}}$, $A_{\mathrm{w}}$, $Q_{\mathrm{c}}$ and
    $Q_{\mathrm{w}}$, derived from \citet{Paterson.Budd.1982} and given in
    \citet[Eqn.~5]{Bueler.Brown.2009}, and $E_{\text{SIA}}=1$.
  \item Our \emph{hard ice} configuration includes rheological parameters,
    $A_{\mathrm{c}}$, $A_{\mathrm{w}}$, $Q_{\mathrm{c}}$ and
    $Q_{\mathrm{w}}$, derived from \citet[p.~72 and
    76]{Cuffey.Paterson.2010}, and $E_{\text{SIA}}=1$, which correspond to
    a~stiffer rheology than that used in the control run.
  \item Our \emph{soft ice} configuration includes rheological parameters
    from \citet{Cuffey.Paterson.2010}, and $E_{\text{SIA}}=5$, the
    recommended value for ice age polar ice
    \citep[p.~77]{Cuffey.Paterson.2010}.
\end{itemize}
%
      An additional simulation using the ice rheology from
      \citet{Cuffey.Paterson.2010} and $E_{\text{SIA}}=2$, the recommended
      value for Holocene polar ice \citep[p.~77]{Cuffey.Paterson.2010} has
      been performed, but is not presented here, since its results were very
      similar to that of our default run.

      Actual parameter values for $A_{\mathrm{c}}$, $A_{\mathrm{w}}$,
      $Q_{\mathrm{c}}$, $Q_{\mathrm{w}}$ and $E_{\text{SIA}}$ used in our
      simulations are given in Table~\ref{tab:sens_params}, while the effect
      of the three different parametrisations on temperature-dependent ice
      softness, $A$, is illustrated in Fig.~\ref{fig:sens_plot_rheo}.

      Surface air
      temperature derived from the climate forcing (Sect.~\ref{sec:atm})
      provides the upper boundary condition to the ice enthalpy
      model. Temperature is computed in the ice and in the bedrock to
      a~depth of 3\,\unit{km} below the ice-bedrock interface, where it is
      conditioned by a~lower boundary geothermal heat flux of
      $q_{\mathrm{G}}=70$\,\unit{mW\,m^{-2}}. Although this uniform value
      does
      not account for the high spatial geothermal variability in the region
      \citep{Blackwell.Richards.2004}, it is, on average, representative of
      available heat flow measurements. In the low-resolution simulations,
      the vertical grid consists of 31~temperature layers in the bedrock and
      up to 51~enthalpy layers in the ice sheet, corresponding to a~vertical
      resolution of 100\,\unit{m}. The high-resolution simulations use
      61~bedrock layers and up to 101~ice layers with a~vertical resolution
      of 50\,\unit{m}.

>>>>

    \idea{In the following, I repeat all parameters in the text and table. Is
          this necessary? Equations are the same as in my Cordillera paper.
          Should I remove them?}

@@

    Ice deformation is governed by the constitutive law for ice
    \citep{Glen.1952, Nye.1953},
%
\begin{align}
&\label{eqn:glenslaw}
&\vec{\dot{\epsilon}} = A\,\tau_{\mathrm{e}}^{n-1}\,\vec{\tau} \,.
\end{align}
%
    where $\vec{\dot{\epsilon}}$ is the strain-rate tensor, $\vec{\tau}$ the
    deviatoric stress tensor, and $\tau_{\mathrm{e}}$ the effective stress
    defined in our case by
    ${\tau_{\mathrm{e}}}^2=\frac{1}{2}\mathrm{tr}(\vec{\tau}^2)$. The ice
    softness coefficient, $A$, depends on ice temperature, $T$, pressure, $p$,
    and water content, $\omega$, through a~piece-wise Arrhenius-type law,
%
\begin{align}
&\label{eqn:softness}
&A = E\cdot
\begin{cases}
A_{\mathrm{c}} \,e^\frac{-Q_{\mathrm{c}}}{RT_{\text{pa}}}            & \text{if}\ T_{\text{pa}}  <  T_{\mathrm{c}} \,, \\
A_{\mathrm{w}} (1+f\omega)\,e^\frac{-Q_{\mathrm{w}}}{RT_{\text{pa}}} & \text{if}\ T_{\text{pa}} \ge T_{\mathrm{c}} \,,
\end{cases}
\end{align}
%
    where $T_{\text{pa}}$ is the pressure-adjusted ice temperature calculated
    using the Clapeyron relation, $T_{\text{pa}}=T-{\beta}p$.
    $R=8.31441$\,\unit{J\,mol^{-1}\,K^{-1}} is the ideal gas constant, and
    $A_{\mathrm{c}}=2.847\e{-13}$\,\unit{Pa^{-3}\,s^{-1}},
    $A_{\mathrm{w}}=2.356\e{-2}$\,\unit{Pa^{-3}\,s^{-1}},
    $Q_{\mathrm{c}}=6.0\e4$\,\unit{J\,mol^{-1}}, and
    $Q_{\mathrm{w}}=11.5\e4$\,\unit{J\,mol^{-1}} are constant parameters
    corresponding to values measured below and above a~critical temperature
    threshold $T_{\mathrm{c}}=-10$\,\unit{{\degree}C}
    \citep[p.~72]{Cuffey.Paterson.2010}. The water fraction, $\omega$, is
    capped at a~maximum value of 0.01, above which no measurements are
    available \citep[Eq.~5.7]{Lliboutry.Duval.1985, Greve.1997}. Finally, $E$
    is a~non-dimensional enhancement factor which can take different values,
    $E_{\text{SIA}}=2$, in the SIA and $E_{\text{SSA}}=1$, in the SSA, as
    recommended for Holocene polar ice \citep[p.~77]{Cuffey.Paterson.2010}.

    In all our simulations, we set constant the power-law exponent, $n=3$,
    according to \citet[p.~55--57]{Cuffey.Paterson.2010}, the Clapeyron
    constant, $\beta=7.9\times 10^{-8}$\,\unit{K\,Pa^{-1}}, according to
    \citet{Luthi.etal.2002}, and the water fraction coefficient, $f=181.25$,
    according to \citet{Lliboutry.Duval.1985}. These fixed parameter values are
    summarized in Table~\ref{tab:params}.

    Surface air temperature derived from the climate forcing
    (Sect.~\ref{sec:atm}) provides the upper boundary condition to the ice
    enthalpy model. Temperature is computed in the ice and in the bedrock to
    a~depth of 3\,\unit{km} below the ice-bedrock interface, where it is
    conditioned by a~lower boundary geothermal heat flux estimate from multiple
    geothermal proxies \citep[similarity method]{Goutorbe.etal.2011}. In the
    2\,km-resolution simulations, the vertical grid consists of 31~temperature
    layers in the bedrock and up to 126~enthalpy layers in the ice,
    corresponding to vertical resolutions of 100 and 40\,\unit{m},
    respectively. The 1\,km-resolution simulation uses 61~bedrock layers and up
    to 251~ice layers with a~vertical resolutions of 50 and 20\,\unit{m}.
